---
title: 'A guide to robust statistical methods in neuroscience: Figure 2'
author: "Rand R. Wilcox & Guillaume A. Rousselet"
date: "`r Sys.Date()`"
output:
  pdf_document:
    fig_caption: no
    number_sections: no
    toc: yes
    toc_depth: 2
---

We perform a simulation of false positives and true positives when sampling from normal and lognormal distributions. The false positive results are illustrated in panel A; the true positive (power) results are illustrated in panel B.

# Dependencies
```{r}
library(ggplot2)
library(cowplot)
library(tidyr)
library(tibble)
# library(beepr) # for simulation only
source("./code/Rallfun-v40.txt")
source("./code/theme_gar.txt")
source("./code/xtrafun.R")
```

## Counter function

Function used to print console updates during the simulation.

```{r, eval=FALSE}
sim.counter <- function(S, nsim, inc){
  if(S == 1){
    # print(paste(nsim,"iterations:",S))
    cat(nsim,"iterations:",S)
    beep(2)
  }
  if(S %% inc == 0){
    # print(paste("iteration",S,"/",nsim))
    cat(" /",S)
    beep(2)
  }
}
```

# Simulation: false positives

Before you commit to 20,000 simulations, try with 1,000, because it might take a long time to compute. The code will output an iteration update every 500 simulations. The simulation includes:  
- multiple sample sizes;  
- sampling from normal and lognormal distributions;  
- t-tests using means and 20% trimmed means;  
- a parametric test of medians.

## Define parameters
```{r}
nsim <- 20000 # number of simulation iterations
nvec <- c(seq(10,150,10), seq(200,400,50), seq(500,800,100)) # vector of sample sizes to test
max.size <- max(nvec)
ES <- 0 # effect size
```

## Run simulation

No need to run this chunk; the results are loaded in the next chunk for convenience. If you run your own simulation, change the name of the file at the end of the chunk, otherwise the results used in the article will be overwritten.  

```{r eval = FALSE}
set.seed(45) # set random number generator for reproducibility

# population values
logn.m <- exp(0.5) # mean of standard lognormal distribution
logn.md <- exp(0) # median of standard lognormal distribution
logn.tm <- ghtrim(g=1) + logn.md # trimmed mean of lognormal distribution

# 6 conditions: normal / skewed x mean / trimmed mean / median
simres <- array(0, dim = c(6, length(nvec), nsim)) 

simsteps <- 500 # get a console update every simsteps iterations

for(S in 1:nsim){ # simulation iterations
  
  sim.counter(S, nsim, simsteps)
  
  large.norm.sample <- rnorm(max.size) # normal sample
  large.lnorm.sample <- rlnorm(max.size) # lognormal sample
  
  for(N in 1:length(nvec)){ # sample sizes
      
      # sub-sample + shift by effect size
      norm.sample <- large.norm.sample[1:nvec[N]] + ES
      lnorm.sample <- large.lnorm.sample[1:nvec[N]] + ES
      
      # normal population
      # t-test on mean
      simres[1,N,S] <- trimci(norm.sample, tr=0, pr=FALSE)$p.value 
      # t-test on 20% trimmed mean
      simres[2,N,S] <- trimci(norm.sample, tr=0.2, pr=FALSE)$p.value
      # median test
      simres[3,N,S] <- sintv2(norm.sample, pr=FALSE)$p.value
      
      # lognormal population
      # t-test on mean
      # subtract mu so the mean is zero on average
      simres[4,N,S] <- trimci(lnorm.sample - logn.m, tr=0, pr=FALSE)$p.value 
      # t-test on 20% trimmed mean
      # subtract population tm so the tm is zero on average
      simres[5,N,S] <- trimci(lnorm.sample - logn.tm, tr=0.2, pr=FALSE)$p.value
      # median test
      # subtract population md so the md is zero on average
      simres[6,N,S] <- sintv2(lnorm.sample - logn.md, pr=FALSE)$p.value
      
    }
}
beep(sound = "mario")
save(simres, file = "./data/simres_fp.RData") 
```

# Plot simulation results

## Get results

We get the type I error rate for each combination of test and sample size.

```{r}
load("./data/simres_fp.RData")
res <- apply(simres <= 0.05, c(1,2), mean)
```

## Figure using base R 

```{r}
plot(nvec, seq(0, 0.2, length.out=length(nvec)), xlab='Sample size', ylab='Type I error rate', type='n')
lines(nvec, res[1,], lty=1, col="orange") # normal: t-test on mean
lines(nvec, res[2,], lty=1, col="green") # normal: t-test on 20% trimmed mean
lines(nvec, res[3,], lty=1, col="blue") # normal: median test
lines(nvec, res[4,], lty=2, col="orange") # lognormal: t-test on mean
lines(nvec, res[5,], lty=2, col="green") # lognormal: t-test on 20% trimmed mean
lines(nvec, res[6,], lty=2, col="blue") # lognormal: median test
abline(0.075, 0, lty=1)
legend(300, 0.2, 
       c("N: mean", "N: tm", "N: md", "LN: mean", "LN: tm", "LN: md"), 
       lty = c(1,1,1,2,2,2),
       col = c("orange","green","blue","orange","green","blue"))
```

## Figure using `ggplot2`

```{r, warning=FALSE}
nvec.labs <- c("10","","","","50","","","","","100","","","","","150","200","250","300","350","400","500","600","700","800")

# create data frame
y <- c(res[1,],res[2,],res[3,],
       res[4,],res[5,],res[6,])
Distribution <- c(rep('Normal',length(nvec)*3), rep('Lognormal',length(nvec)*3))
Estimator <- c(rep('Mean',length(nvec)), rep('Trimmed mean',length(nvec)), rep('Median',length(nvec)), 
               rep('Mean',length(nvec)), rep('Trimmed mean',length(nvec)), rep('Median',length(nvec)))

df <- tibble(x = rep(nvec, 6), # sample size
             y = y, # `Type I error probability
             Distribution = as.factor(Distribution), 
             Estimator = as.factor(Estimator))

df$Estimator <- keeporder(df$Estimator)
df$Distribution <- keeporder(df$Distribution)

# make plot
pA <- ggplot(df, aes(x, y)) + theme_gar +
  # Bradley's (1978) satisfactory range
  geom_ribbon(aes(ymin = 0.025, ymax = 0.075), fill = "grey80") +
  # Bradley's (1978) ideal range
  geom_ribbon(aes(ymin = 0.045, ymax = 0.055), fill = "grey90") +
  geom_abline(intercept = 0.05, slope = 0) + # 0.05 reference line
  geom_line(aes(linetype=Distribution, colour=Estimator), linewidth=0.75) +
  scale_colour_manual(values = c("#56B4E9", "#D55E00", "black")) + ##009E73
  theme(axis.title.x = element_text(size = 18),
        axis.text.x = element_text(size = 14),
        axis.text.y = element_text(size = 16),
        axis.title.y = element_text(size = 18),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        legend.position = c(0.55, 0.8),
        legend.box = "horizontal") +
  scale_x_continuous(breaks = nvec, labels = nvec.labs) +
  scale_y_continuous(limits = c(0,0.16), 
                     breaks = seq(0, 0.15, 0.025)) +
  coord_cartesian(ylim = c(0.02, 0.130)) +
  labs(x = "Sample size", y = "Type I error probability")
pA
```

## Get values reported in the text

```{r, warning=FALSE}
res1 <- round(approx(x=res[4,], y=nvec, xout=0.075)$y)
res2 <- round(approx(x=res[4,], y=nvec, xout=0.055)$y)
```

### T-test on means

False positive probabilities when sampling from a lognormal population:

- For n=30: `r round(res[4,nvec==30], digits=3)`;  
- For n=100: `r round(res[4,nvec==100], digits=3)`;  
- To reach 0.075, n should be at least: `r res1`;  
- To reach 0.055, n should be at least: `r res2`;  
- For n=600: `r round(res[4,nvec==600], digits=3)`;  
- For n=700: `r round(res[4,nvec==700], digits=3)`.  

### Inference on 20% trimmed means

False positive probabilities when sampling from a lognormal population:

- For n=30: `r round(res[5,nvec==30], digits=3)`;  
- For n=100: `r round(res[5,nvec==100], digits=3)`.  

False positive probabilities when sampling from a normal population:

- For n=10: `r round(res[2,nvec==10], digits=3)`;  
- For n=20: `r round(res[2,nvec==20], digits=3)`.  

### Inference on medians

False positive probabilities when sampling from a lognormal population:

- For n=10: `r round(res[6,nvec==10], digits=3)`;  
- For n=30: `r round(res[6,nvec==30], digits=3)`;  
- For n=100: `r round(res[6,nvec==100], digits=3)`.  

False positive probabilities when sampling from a normal population:

- For n=10: `r round(res[3,nvec==10], digits=3)`;  
- For n=20: `r round(res[3,nvec==20], digits=3)`.  

# Simulation: true positives

## Define parameters
```{r}
nsim <- 20000 # number of simulation iterations
nvec <- c(seq(10,100,10), seq(125,250,25), 300) # vector of sample sizes to test
max.size <- max(nvec)
ES <- 0.5 # effect size
```

## Run simulation

No need to run this chunk; the results are loaded in the next chunk for convenience. If you run your own simulation, change the name of the file at the end of the chunk, otherwise the results used in the article will be overwritten.  

```{r eval=FALSE}
set.seed(45) # set random number generator for reproducibility

# population values
logn.m <- exp(0.5) # mean of standard lognormal distribution
logn.md <- exp(0) # median of standard lognormal distribution
logn.tm <- ghtrim(g=1) + logn.md # trimmed mean of lognormal distribution

# 6 conditions: normal / skewed x mean / trimmed mean / median
simres <- array(0, dim = c(6, length(nvec), nsim)) 

simsteps <- 500 # get a console update every simsteps iterations

for(S in 1:nsim){ # simulation iterations
  
  sim.counter(S, nsim, simsteps)
  
  large.norm.sample <- rnorm(max.size) # normal sample
  large.lnorm.sample <- rlnorm(max.size) # lognormal sample
  
  for(N in 1:length(nvec)){ # sample sizes
      
      # sub-sample + shift by effect size
      norm.sample <- large.norm.sample[1:nvec[N]] + ES
      lnorm.sample <- large.lnorm.sample[1:nvec[N]] + ES
      
      # normal population
      # t-test on mean
      simres[1,N,S] <- trimci(norm.sample, tr=0, pr=FALSE)$p.value 
      # t-test on 20% trimmed mean
      simres[2,N,S] <- trimci(norm.sample, tr=0.2, pr=FALSE)$p.value
      # median test
      simres[3,N,S] <- sintv2(norm.sample, pr=FALSE)$p.value
      
      # lognormal population
      # t-test on mean
      # subtract mu so the mean is zero on average
      simres[4,N,S] <- trimci(lnorm.sample - logn.m, tr=0, pr=FALSE)$p.value 
      # t-test on 20% trimmed mean
      # subtract population tm so the tm is zero on average
      simres[5,N,S] <- trimci(lnorm.sample - logn.tm, tr=0.2, pr=FALSE)$p.value
      # median test
      # subtract population md so the md is zero on average
      simres[6,N,S] <- sintv2(lnorm.sample - logn.md, pr=FALSE)$p.value
      
    }
}
beep(sound = "mario")
save(simres, file = "./data/simres_tp.RData") 
```

# Plot simulation results

## Get results

We get the power for each combination of test and sample size.

```{r}
load("./data/simres_tp.RData")
res <- apply(simres <= 0.05, c(1,2), mean)
```

## Base R figure

```{r}
plot(nvec, seq(0, 1, length.out=length(nvec)), xlab='Sample size', ylab='Power', type='n')
lines(nvec, res[1,], lty=1, col="orange") # normal: t-test on mean
lines(nvec, res[2,], lty=1, col="green") # normal: t-test on 20% trimmed mean
lines(nvec, res[3,], lty=1, col="blue") # normal: median test
lines(nvec, res[4,], lty=2, col="orange") # lognormal: t-test on mean
lines(nvec, res[5,], lty=2, col="green") # lognormal: t-test on 20% trimmed mean
lines(nvec, res[6,], lty=2, col="blue") # lognormal: median test
abline(0.8, 0, lty=2)
legend(200, 0.5, 
       c("N: mean", "N: tm", "N: md", "LN: mean", "LN: tm", "LN: md"), 
       lty = c(1,1,1,2,2,2),
       col = c("orange","green","blue","orange","green","blue"))
```


# `ggplot2` figure

```{r}
nvec.labs <- c("10","","","","50","","","","","100","","150","","200","","250","300")

# create data frame
y <- c(res[1,],res[2,],res[3,],
       res[4,],res[5,],res[6,])
Distribution <- c(rep('Normal',length(nvec)*3), rep('Lognormal',length(nvec)*3))
Estimator <- c(rep('Mean',length(nvec)), rep('Trimmed Mean',length(nvec)), rep('Median',length(nvec)), 
               rep('Mean',length(nvec)), rep('Trimmed Mean',length(nvec)), rep('Median',length(nvec)))

df <- tibble(x = rep(nvec, 6), # sample size
             y = y, # `Type I error probability
             Distribution = as.factor(Distribution), 
             Estimator = as.factor(Estimator))

df$Estimator <- keeporder(df$Estimator)
df$Distribution <- keeporder(df$Distribution)

# make plot
pB <- ggplot(df, aes(x, y)) + theme_gar +
  geom_abline(intercept = 0.80, slope = 0, linewidth=0.25) + # 0.80 reference line
  geom_abline(intercept = 0.90, slope = 0, linewidth=0.25) + # 0.90 reference line
  geom_line(aes(linetype=Distribution, colour=Estimator), linewidth=0.75, show.legend = FALSE) +
  scale_colour_manual(values = c("#56B4E9", "#D55E00", "black")) + #009E73
  theme(axis.title.x = element_text(size = 18),
        axis.text = element_text(size = 16),
        axis.title.y = element_text(size = 18),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        legend.position = c(0.65, 0.2),
        legend.box = "horizontal") +
  scale_y_continuous(limits = c(0,1), 
                     breaks = seq(0, 1, 0.2),
                     minor_breaks = seq(0.1,0.9,0.2)) +
  scale_x_continuous(limits = c(0, 300),
                     breaks = nvec, labels = nvec.labs) +
  labs(x = "Sample size", y = "Power (true positive probability)")
pB
```


## combine panels into one figure
```{r, eval=FALSE}
cowplot::plot_grid(pA, pB,
  labels=c("A", "B"), 
  ncol = 1, 
  nrow = 2,
  rel_widths = c(1, 1), 
  label_size = 20, 
  hjust = -0.5, 
  scale=.95,
  align = "h")

# save figure
ggsave(filename='./figures/figure2.pdf',width=7,height=10) 
```

