---
title: 'A guide to robust statistical methods in neuroscience: Figure 8'
author: "Rand R. Wilcox & Guillaume A. Rousselet"
date: "`r Sys.Date()`"
output:
  pdf_document:
    fig_caption: no
    number_sections: no
    toc: yes
    toc_depth: 2
---

**Development of Anxiety-Like Behaviors after Mild Traumatic Brain Injury**  
Almeida-Suhett et al. (2014)

The experimental groups  (5-6 week old male, Sprague--Dawley rats) received a mild controlled cortical impact (CCI). The outcome measure was stereologically estimated total number of GAD-67-positive cells in the basolateral amygdala (BLA). There are three independent groups: sham-treated controls that received a craniotomy, but no CCI injury, measures taken 1 day after CCI, and measures taken 7 days later.

# Dependencies
```{r}
library(ggplot2)
library(tibble)
library(tidyr)
source("./code/Rallfun-v40.txt")
source("./code/theme_gar.txt")
```

# Load data
```{r}
# First six columns deal with neurons, final six with GAD-67
prager=read.table('./data/Prager_Behvior_data.tex',skip=1,header=T)
```

# base R ipsi boxplots
```{r}
boxplot(prager[,c(7,9,11)])
```

# base R contra boxplots
```{r}
boxplot(prager[,c(8,10,12)])
```
# Make `ggplot2 figure`

## Load data and reformat
```{r, warning=FALSE}
# ipsi <- prager[,c(7,9,11)]
# contra <- prager[,c(8,10,12)]
df <- as_tibble(prager[,7:12])
df <- gather(df, group, BLA, na.rm=TRUE)
df <- separate(df, group, c("location", "condition"), sep = "\\.")
df$BLA <- as.numeric(df$BLA)
df$location <- as.factor(df$location)
df$condition <- factor(df$condition, levels = c("S", "1", "7"))
```

## Create figure
```{r}
ggplot(df, aes(condition, BLA)) + theme_gar + 
  geom_jitter(colour = "grey", width = 0.05) + 
  geom_boxplot(outlier.colour = "grey70", outlier.shape = 16,
               outlier.size = 3, outlier.alpha = 0, size = 0.25, alpha = 0) +
  stat_summary(fun = "mean",  geom = "point", size = 1.5) +
  theme(axis.text.x = element_text(size = 16),
        axis.text.y = element_text(size = 16),
        axis.title.x = element_text(size = 18),
        axis.title.y = element_text(size = 18),
        plot.title = element_text(size = 20),
        legend.position="none") +
  scale_x_discrete(breaks = c("S", "1", "7"),
                   labels = c("Sham", "Day 1", "Day 7")) +
  labs(x = "Conditions",
       y = "number of GAD-67-positive cells\nin the basolateral amygdala") +
  scale_y_continuous(limits = c(2500, 8000), breaks = seq(3000, 8000, 1000)) +
  facet_grid(. ~ location, 
             labeller = labeller(group = c(Ipsi = "Ipsilateral", 
                                           Contra = "Contralateral"))) +
  theme(strip.text.x = element_text(size = 20, colour = "white"),
        strip.background = element_rect(colour="darkgrey", fill="darkgrey"))

# save figure
ggsave(filename='./figures/figure8.pdf',width=6,height=6) 
```

# Comparisons: ipsilateral conditions

## Standard t-test

### Day 7 vs control (sham)
```{r}
t.test(prager[,7],prager[,9],var.equal =T)  
```
### Day 1 vs control (sham)
```{r}
t.test(prager[,7],prager[,11],var.equal =T) 
```

So by Bonferroni, fail to reject  sham vs. Day 1. 
But reject sham vs Day 7. Note, however, that for Day 1, the 
improvement on the Bonferroni method derived by Hochberg (1988) rejects.

```{r}
p.adjust(c(0.0375, 0.0001406), method = "hochberg")
```

Now repeat this using a 20\% trimmed mean using a non-bootstrap
method followed by a bootstrap method, then compare medians,
then Cliff's improvement of WMW, next compare medians
using the Harrell--Davis estimator, followed by the symmetry test, and 
finally use the median of D=X-Y.

Again focus on Ipsi and day 1 vs. sham. 

## Ipsi: Sham vs. Day 1

### Non-bootstrap 20% trimmed mean
```{r}
yuen(prager[,7],prager[,11])  
```

### 20% trimmed mean + bootstrap
```{r}
set.seed(44)
nboot <- 2000
trimpb2(prager[,7],prager[,11], nboot=nboot)  
```

In this situation, with relatively small  sample sizes, the expectation is that trimpb2 (bootstrap method) is better than yuen. As illustrated here, the choice between these two methods can make a practical difference. 

### medians + bootstrap
```{r}
set.seed(44)
medpb2(prager[,7],prager[,11], nboot=nboot)  
```

### Improved Wilcoxon-Mann-Whitney
```{r}
cidv2(prager[,7],prager[,11]) 
```

### Harrell-Davis estimator + bootstrap
```{r}
set.seed(44)
pb2genMC(prager[,7],prager[,11],est=hd, nboot=nboot)
```

### Symmetry test based on D=X-Y
```{r}
set.seed(44)
cbmhd(prager[,7],prager[,11], nboot=nboot) 
```

If the two distributions are identical, the distribution of $D=X-Y$ 
should be symmetric about zero, so the sum of the lower and upper
quartiles should be zero. The symmetry test performed by `cbmhd` indicates that this is not the
case. Even the lower quartile is estimated to be greater than zero.

### Median of the typical difference between randomly sampled observations from each group
That is,focus on D=X-Y rather than the median of X and Y.
```{r}
set.seed(44)
wmwpb(prager[,7],prager[,11], nboot=nboot)
```

Did not reject with medians, $p=0.086$, but based on the median
of typical difference, $p=0.021$, merely illustrating that the point of
view can matter. 

### Shift function (compare deciles)
Comparing quantiles via qcomhd, the 0.75 and 0.9 quantiles differ ($p=0.40$ and $p< 0.001$),
otherwise no significant differences.

```{r}
set.seed(44)
qcomhd(prager[,7],prager[,11], nboot=nboot)
```

## Ipsi: Sham vs. Day 7

### Non-bootstrap 20% trimmed mean
```{r}
yuen(prager[,7],prager[,9])  
```

### 20% trimmed mean + bootstrap
```{r}
set.seed(44)
nboot <- 2000
trimpb2(prager[,7],prager[,9], nboot=nboot)  
```

### medians + bootstrap
```{r}
set.seed(44)
medpb2(prager[,7],prager[,9], nboot=nboot)  
```

### Improved Wilcoxon-Mann-Whitney
```{r}
cidv2(prager[,7],prager[,9]) 
```

### Harrell-Davis estimator + bootstrap
```{r}
set.seed(44)
pb2genMC(prager[,7],prager[,9],est=hd, nboot=nboot)
```

### Symmetry test based on D=X-Y
```{r}
set.seed(44)
cbmhd(prager[,7],prager[,9], nboot=nboot) 
```

### Median of the typical difference between randomly sampled observations from each group
```{r}
set.seed(44)
wmwpb(prager[,7],prager[,9], nboot=nboot)
```

### Shift function (compare deciles)
```{r}
set.seed(44)
qcomhd(prager[,7],prager[,9], nboot=nboot)
```


# Comparisons: contralateral conditions

Repeat the above using contra.

First do T-tests:

```{r}
t.test(prager[,8],prager[,10],var.equal =T)  # Day 7 vs control (sham)
t.test(prager[,8],prager[,12],var.equal =T) 
```

So based on Student's T, sham vs. Day 1, not significant, sham vs. Day 7 is.

Now look at sham vs. Day 7 using robust methods.

## Contra: Sham vs. Day 7

### Non-bootstrap 20% trimmed mean
```{r}
yuen(prager[,8],prager[,10])  
```

### Bootstrap + trimmed means
```{r}
set.seed(44)
trimpb2(prager[,8],prager[,10], nboot = nboot)
```

### Medians
```{r}
set.seed(44)
medpb2(prager[,8],prager[,10], nboot = nboot)  
```

### Cliff
```{r}
cidv2(prager[,8],prager[,10]) 
```

### Harrell-Davis estimator
```{r}
set.seed(44)
pb2genMC(prager[,8],prager[,10],est=hd, nboot = nboot) 
```

### Symmetry test based on D=X-Y
```{r}
cbmhd(prager[,8],prager[,10], nboot = nboot) 
```

### Median of the typical difference
```{r}
set.seed(44)
wmwpb(prager[,8],prager[,10], nboot = nboot) 
```

So even setting FWE (Familywise error, the probability of one or more
Type I errors) equal to 0.001, all of the robust methods reject.

## Contra: sham versus Day 1

### Non-bootstrap 20% trimmed mean
```{r}
yuen(prager[,8],prager[,12])  
```

### Bootstrap + trimmed means
```{r}
set.seed(44)
trimpb2(prager[,8],prager[,12], nboot = nboot)
```

### Medians
```{r}
set.seed(44)
medpb2(prager[,8],prager[,12], nboot = nboot)  
```

### Cliff
```{r}
cidv2(prager[,8],prager[,12]) 
```

### Harrell-Davis estimator
```{r}
set.seed(44)
pb2genMC(prager[,8],prager[,12],est=hd, nboot = nboot) 
```

### Symmetry test based on D=X-Y
```{r}
set.seed(44)
cbmhd(prager[,8],prager[,12], nboot = nboot) 
```

### Median of the typical difference
```{r}
set.seed(44)
wmwpb(prager[,8],prager[,12], nboot = nboot) 
```

So for contra, sham versus day 1, no significant differences.



