---
title: 'A guide to robust statistical methods in neuroscience: Figure 6'
author: "Rand R. Wilcox & Guillaume A. Rousselet"
date: "`r Sys.Date()`"
output:
  pdf_document:
    fig_caption: no
    number_sections: no
    toc: yes
    toc_depth: 2
---

Mancini et al. (2014) report results aimed at providing a
whole-body mapping of spatial acuity for pain. Here, we provide additional results
based on the data from their second experiment. The sample size is $n=10$.  Briefly, spatial acuity was assessed by measuring 2-point discrimination (2PD) thresholds for both pain and touch in 11 body territories.

The data can also be downloaded here:

Mancini, Flavia (2016): ANA_mancini14_data.zip. figshare.
https://doi.org/10.6084/m9.figshare.3427766.v1

# Dependencies
```{r}
library(ggplot2)
library(cowplot)
library(tibble)
source("./code/Rallfun-v40.txt")
source("./code/theme_gar.txt")
source("./code/xtrafun.R")
```

# Make figure

## get data
```{r}
Man = read.table(file='./data/Man_try.tex',skip=1,header=T)
```

## make data frame 1: marginal distributions -------------------------------------
```{r}
DT <- Man[,13:18] # discrimination thresholds
colnames(DT) <- c('FH', 'S', 'FA', 'H', 'B', 'T')
Participants <- seq(1,nrow(DT))
df <- as_tibble(cbind(DT,Participants))
df$Participants <- as.factor(df$Participants)
df <- tidyr::gather(df, Positions, Thresholds, -Participants) # make long format
df$Positions <- as.factor(df$Positions)
df$Positions <- keeporder(df$Positions)
dfA <- df
```

## data frame of summary statistics 

### medians

Summarise each column using the Harrell-Davis estimate of the 50th quantile.

```{r}
tmp <- as_tibble(DT)
df.hd <- enframe(apply(tmp, 2, hd), name="Positions", value="Thresholds")
df.hd$Positions <- as.factor(df.hd$Positions)
```

Use the median instead

```{r}
tmp <- as_tibble(DT)
df.md <- enframe(apply(tmp, 2, median), name="Positions", value="Thresholds")
df.md$Positions <- as.factor(df.md$Positions)
```

### means

```{r}
tmp <- as_tibble(DT)
df.mn <- enframe(apply(tmp, 2, mean, trim = 0), name="Positions", value="Thresholds")
df.mn$Positions <- as.factor(df.mn$Positions)
```

## parameters for plots
```{r}
axis_size <- 14
axis_title_size <- 16
```

## panel A: marginals
```{r}
pA <- ggplot(dfA, aes(x=Positions, y=Thresholds)) + theme_gar +
  geom_line(size = 0.5, aes(group = Participants, colour = Participants),
            alpha = .6) +
  geom_point(aes(fill = Participants), shape = 21, size = 3, 
             colour = "black", alpha = .6) +
  theme(axis.text = element_text(size = axis_size),
        axis.title = element_text(size = axis_title_size),
        legend.position = "none") +
  scale_y_continuous(limits = c(0,3.5), breaks = seq(0,3.5,0.5)) +
  xlab("Positions") +
  # superimpose medians
  geom_line(data = df.md, aes(group = 1), linewidth = 1.2) +
  geom_point(data = df.md, aes(group = 1), size = 3.5)
  # # superimpose means
  # geom_line(data = df.mn, aes(group = 1), linewidth = 1, colour = "grey") +
  # geom_point(data = df.mn, aes(group = 1), size = 2, colour = "grey")
pA
```

## make data frame 2: difference distributions -----------------------------------
```{r}
# compute all pairwise differences
J <- ncol(DT)
n <- nrow(DT)
Jm <- J-1
d <- (J^2-J)/2
con <- matrix(0,J,d)
newnames <- vector(mode = "character", length = d)
cnames <- c('FH', 'S', 'FA', 'H', 'B', 'T')
id<-0
for (j in 1:Jm){
  jp <- j+1
  for (k in jp:J){
    id<-id+1
    con[j,id] <- 1
    con[k,id] <- 0-1
    newnames[id] <- paste(cnames[j],'-',cnames[k], sep = "")
  }
}
xx <- matl(DT)%*%con

# make data frame
colnames(xx) <- newnames
df <- as_tibble(xx)
df <- tidyr::gather(df, Contrasts, Differences) # make long format
df$Contrasts <- as.factor(df$Contrasts)
df$Contrasts <- keeporder(df$Contrasts)
dfB <- df

# parameters for plots
axis_title_size <- 16
```


# panel B: differences
```{r}
pB <- ggplot(dfB, aes(x=Contrasts, y=Differences, 
                      colour = Contrasts, fill = Contrasts)) +
  theme_gar + 
  geom_point(shape = 21, size = 3, colour = "black") +
  geom_abline(intercept = 0, slope = 0, linetype = 2) +
  theme(axis.text.y = element_text(size = 14),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 12),
        axis.title = element_text(size = axis_title_size),
        legend.position = "none") +
  xlab("Contrasts") +
  # add median
  stat_summary(fun = median, fun.min = median, fun.max = median,
               geom = "crossbar", 
               width = 0.5, 
               colour = "black", 
               size = 0.5)
  # # add mean
  # stat_summary(fun = mean, fun.min = mean, fun.max = mean,
  #              geom = "crossbar", 
  #              width = 0.5, 
  #              colour = "grey", 
  #              size = 0.3) 
pB
```

## combine panels into one figure - AB only
```{r, eval=FALSE}
cowplot::plot_grid(pA, pB,
  labels=c("A", "B"), 
  ncol = 1, 
  nrow = 2,
  rel_widths = c(1, 1), 
  label_size = 20, 
  hjust = -0.5, 
  scale=.95,
  align = "h")

# save figure
# ggsave(filename='./figures/figure6.pdf',width=10,height=10) 
```

## Reject at alpha level for different methods?

```{r}
alpha.val <- 0.05
rej.res <- matrix(data = NA, nrow = 4, ncol = d)
diffs <- matrix(data = NA, nrow = 4, ncol = d)

# Means -- no correction for multiple comparisons
res <- wmcp(Man[,13:18],tr=0)  # 10 sig
rej.res[1,] <- res$test[,4] < alpha.val
diffs[1,] <- res$psihat[,3]

# Means + Hochberg
rej.res[2,] <- res$test[,4] < res$test[,5]
diffs[2,] <- res$psihat[,3]

# 20% trimmed means + Hochberg
set.seed(666)
res <- wmcppb(Man[,13:18], nboot=nboot, tr=tr) # 12 sig
rej.res[3,] <- res$output[,3] < res$output[,4]
diffs[3,] <- res$output[,2]

# Medians + Hochberg
set.seed(666)
res <- dmedpb(Man[,13:18], nboot=nboot) # 10 sig
rej.res[4,] <- res$output[,3] < res$output[,4]
diffs[4,] <- res$output[,2]
```

### Illustrate results

Size of disks reflects size of differences.

```{r}
dfC <- tibble(Reject = as.factor(as.vector(rej.res)),
             Methods = as.factor(rep(c("M(nc)", "M", "TM", "MD"), d)),
             Contrasts = as.factor(rep(newnames, each = 4)),
             Differences = as.vector(abs(diffs)) # use absolute differences
             )

dfC$Methods <- keeporder(dfC$Methods) 
dfC$Contrasts <- keeporder(dfC$Contrasts)

pC <- ggplot(dfC, aes(x = Contrasts, y = Methods, 
                      colour = Reject, fill = Reject, size = Differences)) +
  theme_gar + 
  geom_point(shape = 21, colour = "black") +
  scale_fill_manual(values = c("white", "darkgrey")) +
  scale_y_discrete(limits=rev) +
  scale_size_continuous(range = c(2,8), guide = "none") +
  theme(axis.text.y = element_text(size = 14),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 12),
        axis.title = element_text(size = axis_title_size),
        legend.position = "bottom") +
  guides(fill = guide_legend(override.aes = list(size = 8)))

pC
```


## combine panels into one figure - ABC
```{r, eval=FALSE}
cowplot::plot_grid(pA, pB, pC,
  labels=c("A", "B", "C"), 
  ncol = 1, 
  nrow = 3,
  rel_widths = c(1, 1, 1),
  rel_heights = c(1, 1, 0.6),
  label_size = 20, 
  hjust = -0.5, 
  scale=.95,
  align = "v")

# save figure
ggsave(filename='./figures/figure6.pdf',width=10,height=15) 
```

# Results reported in the text

## Comparing foot and thigh pain measures

### T-test
```{r}
t.test(Man$FootP-Man$thighP)
```

### Method using 20% trimmed means
```{r}
set.seed(666)
nboot <- 2000 # number of bootstrap samples
tr <- 0.2 # amount of trimming
trimpb(Man$FootP-Man$thighP, nboot=nboot, tr=tr)
```

### Method using medians
```{r}
set.seed(666)
dmedpb(Man$FootP,Man$thighP,nboot=nboot)
```

## All pairwise comparisons

All pairwise comparisons, based on touch, were performed for the following body parts: forehead, shoulder,  forearm, palm, dorsum, back and  thigh. The probability of one or more Type I errors was controlled using an improvement on the Bonferroni method derived by Hochberg (1988).

### Means
```{r}
wmcp(Man[,13:18],tr=0)  # 10 sig
```

### Medians
```{r}
set.seed(666)
dmedpb(Man[,13:18], nboot=nboot) # 10 sig
```

### Trimmed means
```{r}
set.seed(666)
wmcppb(Man[,13:18], nboot=nboot, tr=tr) # 12 sig
```

### Sign test
```{r, eval=FALSE}
signmcp(Man[,13:18], method='SD') 
```

